---
title: "MATH 18/19: Overall analysis"
author: "George Kinnear"
date: "28/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(tidyverse)
library(knitr)
library(BEST) # for Bayesian estimation in place of t-tests
```

## Read data

Read in the `data_math-1819.csv` file which contains the codes assigned to all items, and produce some summaries of this.

```{r read-data, message=FALSE, warning=FALSE, include=FALSE}
source(file = here::here("script_read-data.R"))

Course_totals = ratings %>%
  group_by(Course) %>%
  summarise(
    total_question = n(),
    total_marks = sum(Marks)
  )

```

Summary of papers in the sample:

```{r summary-papers, echo=FALSE}
MATH_Group_proportions %>%
  group_by(Course, Year) %>%
  tally() %>% 
  group_by(Course) %>% 
  summarise(
    num_papers = n_distinct(Year),
    papers = paste0(Year, collapse = ", ")
  ) %>% 
  left_join(Course_totals) %>% 
  kable(booktabs = TRUE)

Course_totals %>%
  ungroup() %>%
  summarise(
    totalqs = sum(total_question),
    totalmarks = sum(total_marks)
  ) %>% 
  kable(booktabs = TRUE)
```



# Overall proportions

```{r plot-proportions}
MATH_Group_proportions %>%
  # reorder the Course factor by max % group A
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    Course = reorder(Course, gpAmarks, max)
  ) %>%
  ggplot(aes(x=1,y=prop_marks, fill=Course)) +
  geom_boxplot(alpha = 0.4) + 
  geom_point(aes(colour=Course),
             position = position_jitterdodge(),
             size = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     breaks = seq(0,1,0.1)) +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  scale_colour_brewer(palette = "RdBu", direction = -1) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black")
  ) +
  labs(
#    title = "Proportion of marks in each MATH Group",
#    subtitle = "Exams are ordered by proportion of Group A marks",
    x = "MATH Group",
    y = "Proportion of marks"
  ) +
  facet_grid( ~ MATH_Group, switch = "x") +
  ggsave("figs/ABC_allpapers_ranking.pdf",width=15,height=10,units="cm",dpi=300)

```


In this version, we sort by the mean % of group A, and show the means with diamonds

```{r plot-props-sorted}
MATH_Group_proportions %>%
  # reorder the Course factor by mean % group A
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    Course = reorder(Course, gpAmarks, mean)
  ) %>%
  ggplot(aes(x=1,y=prop_marks, fill=Course)) +
  geom_boxplot(alpha = 0.3) + 
  stat_summary(fun.y = "mean",
               geom = "point",
               shape = 23,
               size = 3,
               aes(fill = Course),
               position = position_dodge(0.8)) +
  geom_point(aes(colour=Course),
             position = position_jitterdodge(),
             size = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     breaks = seq(0,1,0.1)) +
  #  scale_fill_brewer(palette = "Set1") +
  #  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  scale_colour_brewer(palette = "RdBu", direction = -1) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black")
  ) +
  labs(
    #    title = "Proportion of marks in each MATH Group",
    #    subtitle = "Exams are ordered by proportion of Group A marks",
    x = "MATH Group",
    y = "Proportion of marks"
  ) +
  facet_grid( ~ MATH_Group, switch = "x") +
  ggsave("figs/ABC_allpapers_boxplots.pdf",width=15,height=10,units="cm",dpi=300)
```

Here we group courses at different education levels

```{r plot-props-grouped}
gpAschooluni = MATH_Group_proportions %>%
  # reorder the Course factor by mean % group A
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    Course = reorder(Course, gpAmarks, mean),
    CourseGroup = case_when(
      Course %in% c("TMUA P1", "TMUA P2", "DiagTest") ~ "Admission",
      Course == "ILA" ~ "University",
      TRUE ~ "School"
    )
  ) %>%
  filter( MATH_Group == "A" ) %>%
  select(Course, Year, Paper, total_marks_for_Group, total_marks_for_paper, prop_marks, CourseGroup)

gpAschooluni %>% 
  ggplot(aes(x=1,y=prop_marks)) +
  geom_boxplot(alpha = 0.3) + 
  stat_summary(fun.y = "mean",
               geom = "point",
               shape = 23,
               size = 3) +
  geom_point(aes(colour=Course),
             size = 1,
             position = position_jitter(0.2)) +
  #  geom_text(aes(label=Course)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     breaks = seq(0,1,0.1),
                     limits = c(0,1)) +
  #  scale_fill_brewer(palette = "Set1") +
  #  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  scale_colour_brewer(palette = "RdBu", direction = -1) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black")
  ) +
  labs(
    #    title = "Proportion of marks in each MATH Group",
    #    subtitle = "Exams are ordered by proportion of Group A marks",
    x = "Exam Level",
    y = "Proportion of marks in MATH Group A"
  ) +
  facet_grid( ~ CourseGroup, switch = "x") +
  ggsave("figs/A_mean_school_vs_uni.pdf",width=15,height=10,units="cm",dpi=300)


```

## Statistical comparisons


### t-test comparison by level

```{r stats-t-test-levels}
gpAschool = gpAschooluni %>% filter(CourseGroup=="School") %>% select(prop_marks) %>% data.matrix
gpA_Alevel = gpAschooluni %>% filter(str_detect(Course, "A-Level")) %>% select(prop_marks) %>% data.matrix
gpAuni = gpAschooluni %>% filter(Course=="ILA") %>% select(prop_marks) %>% data.matrix

t.test(gpAschool, gpAuni)
t.test(gpA_Alevel, gpAuni)
```

### Bayesian comparison

```{r stats-bayes-levels}
BESTout = BESTmcmc(gpAschool,gpAuni)
BESTout
summary(BESTout)
```

This gives the mean proportion of group A:

* for school exams, `r formathdi(summary(BESTout),"mu1",2)`
* for university exams, `r formathdi(summary(BESTout),"mu2",2)`

The difference between these is `r formathdi(summary(BESTout),"muDiff",2)`.

Here are some more diagostic plots for this summary:

```{r stats-bayes-levels-detail}
plot(BESTout)
plot(BESTout, "sd")
plotPostPred(BESTout)
plotAll(BESTout, credMass=0.95, ROPEm=c(-0.1,0.1),
        ROPEeff=c(-0.2,0.2), compValm=0.5)
pairs(BESTout)
```

Looking at university versus A-Level in particular:

```{r stats-bayes-alevel-vs-uni}
BESTout_Alevel = BESTmcmc(gpA_Alevel,gpAuni)
summary(BESTout_Alevel)
plotAll(BESTout_Alevel, credMass=0.95, ROPEm=c(-0.1,0.1),
        ROPEeff=c(-0.2,0.2), compValm=0.5)
```

# Details of each paper

Stacked bars of ABC proprtion by exam paper

```{r plot-abc-stacked-bars}
math_stacked_bars <- function(course_name) {
  MATH_Group_proportions %>%
    filter(Course == course_name) %>%
    # compute the positions of the text labels
    # see http://t-redactyl.io/blog/2016/01/creating-plots-in-r-using-ggplot2-part-4-stacked-bar-plots.html
    ddply(.(Course,Exam), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
    ggplot(aes(x = Exam,
               y = prop_marks,
               fill = fct_rev(MATH_Group))) +
    geom_bar(stat="identity", alpha = 0.9) +
    geom_text(aes(x = Exam, y = 1-pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
              size=4) +
    scale_y_continuous(labels = scales::percent,
                       breaks = seq(0,1,0.1)) +
    scale_fill_manual(values = palMATH) + 
    theme_light(base_size = 12)+
    theme(
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 60, hjust = 1)
    ) +
    labs(
      title = paste(course_name, "by year"),
      fill = "MATH Group",
      x = "Exam",
      y = "Proportion of marks"
    )
}
math_stacked_bars("ILA")
math_stacked_bars("SQA AH")
math_stacked_bars("IB HL")
#math_stacked_bars("A-Level C1-C4")
#math_stacked_bars("A-Level FP3")
```

Do them all in one big plot, using facets:

```{r plot-abc-props-facets, fig.asp=1.2}
MATH_Group_proportions %>%
  filter(!Course %in% c("DiagTest", "TMUA")) %>%
  ddply(.(Course,Exam), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Exam,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity", alpha = 0.9) +
  geom_text(aes(x = Exam, y = 1-pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=2) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) + 
  theme_light(base_size = 10)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = rel(0.7), angle = 30, hjust = 1),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black"),
    legend.position="bottom"
  ) +
  labs(
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  facet_wrap( ~ Course, ncol = 2, scales = "free") +
  ggsave("figs/ABC_allpapers_facets.pdf",width=15,height=20,units="cm",dpi=300)
```




A version of Darlington 2015 Fig 2/3:

```{r plot-abc-school-vs-uni}
MATH_Group_proportions %>%
  filter(!Course %in% c("TMUA", "DiagTest")) %>%
  ungroup() %>%
  group_by(Course,MATH_Group) %>%
  summarise(
    prop_marks = mean(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
#    Course = paste(Course,"\n(n=",n,")"),
    Course = (reorder(Course, gpAmarks, max))
  ) %>%
  ddply(.(Course), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Course,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity", alpha = 0.9) +
  geom_text(aes(x = Course, y = pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=4) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) + 
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(
#    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  ggsave("figs/ABC_school_and_ILA_means.pdf",width=15,height=10,units="cm",dpi=300)

```


Try using facets to show school vs university:

```{r plot-school-vs-uni-facets}
MATH_Group_proportions %>%
  ungroup() %>%
  group_by(Course,MATH_Group) %>%
  summarise(
    prop_marks = mean(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    #    Course = paste(Course,"\n(n=",n,")"),
    Course = (reorder(Course, gpAmarks, max)),
    CourseGroup = if_else(Course %in% c("TMUA", "DiagTest", "ILA"), "University", "School")
  ) %>%
  ddply(.(Course), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Course,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity", alpha = 0.9) +
  geom_text(aes(x = Course, y = pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=4) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) +
  facet_wrap( ~ CourseGroup, ncol = 2, scales = "free") +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black")
  ) +
  labs(
    #    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  ggsave("figs/ABC_school_vs_other_means.pdf",width=15,height=10,units="cm",dpi=300)
```

Showing the school/uni distinction using opacity:

```{r plot-school-vs-uni-alpha}
MATH_Group_proportions %>%
  ungroup() %>%
  group_by(Course,MATH_Group) %>%
  summarise(
    prop_marks = mean(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    #    Course = paste(Course,"\n(n=",n,")"),
    Course = (reorder(Course, gpAmarks, max)),
    CourseGroup = if_else(Course %in% c("TMUA P1", "TMUA P2", "DiagTest", "ILA"), "University", "School")
  ) %>%
  ddply(.(Course), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Course,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity",
           aes(alpha = CourseGroup)) +
  geom_text(aes(x = Course, y = pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=4) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) +
  scale_alpha_manual("Exam Level", values = c("School" = 0.9, "University" = 0.4)) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(
    #    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  ggsave("figs/ABC_all_means.pdf",width=15,height=10,units="cm",dpi=300)
```

Focusing on school exams:

```{r plot-school-means}
MATH_Group_proportions %>%
  filter(!Course %in% c("TMUA", "DiagTest", "ILA")) %>%
  ungroup() %>%
  group_by(Course,MATH_Group) %>%
  summarise(
    prop_marks = mean(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    #    Course = paste(Course,"\n(n=",n,")"),
    Course = (reorder(Course, gpAmarks, max))
  ) %>%
  ddply(.(Course), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Course,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity", alpha = 0.9) +
  geom_text(aes(x = Course, y = pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=4) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) + 
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(
    #    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  ggsave("figs/ABC_school_means.pdf",width=15,height=10,units="cm",dpi=300)
```

# Detailed MATH categories

Like Darlington (2014) Fig 1 - looking in detail at the MATH categories.

```{r tab-math-cats-detail}
MATH_Cat_proportions = ratings %>%
  group_by(Course,Year,Paper,MATH_Group,MATH) %>%
  summarise(
    total_marks_for_Cat = sum(Marks)
  ) %>%
  ungroup() %>%
  # add in the missing "0%" entries
  complete(nesting(Course,Year,Paper),MATH, fill = list(total_marks_for_Cat = 0)) %>%
  mutate(
    # restore the MATH_Group for the "0%" entries
    MATH_Group = str_sub(MATH,1,1)
  ) %>%
  left_join(MATH_Group_proportions) %>%
  mutate (
    prop_marks_cat = total_marks_for_Cat / total_marks_for_Group
  ) %>%
  ungroup()

MATH_Cat_props_by_Course = MATH_Cat_proportions %>%
  group_by(Course, MATH) %>%
  summarise(
    total_Cat = sum(total_marks_for_Cat),
    total_Group = sum(total_marks_for_Group),
    prop_of_Group = total_Cat / total_Group * 100
  )

MATH_Cat_props_by_Course %>%
  select(-total_Cat,-total_Group) %>%
  spread(MATH, prop_of_Group) %>%
  knitr::kable(booktabs = T, digits = 0)
  
```


```{r plot-math-cats-detail}
MATH_Cat_props_by_Course %>%
  filter( str_sub(MATH,1,1) == "A") %>%
  ungroup() %>%
  mutate(
    Course = (reorder(Course, prop_of_Group, max)),
    CourseGroup = if_else(Course %in% c("TMUA", "DiagTest", "ILA"), "University", "School"),
    MATH = fct_recode(MATH, FKFS = "A1", COMP = "A2", RUOP = "A3"),
    prop_of_Group = prop_of_Group / 100
  ) %>%
  ddply(.(Course), transform, pos = cumsum(prop_of_Group) - (0.5 * prop_of_Group)) %>%
  ggplot(aes(x = Course,
             y = prop_of_Group,
             fill = fct_rev(MATH),
             alpha = CourseGroup)) +
  geom_bar(stat="identity") +
  geom_text(aes(x = Course,
                y = pos,
                label = if_else(prop_of_Group <0.01,
                                "",
                                paste0(sprintf("%.0f",100*prop_of_Group),"%"))
                ),
            size = 4,
            alpha = 1) +
  scale_alpha_manual("Exam Level", values = c("School" = 0.9, "University" = 0.5)) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = rev(c(palMATH[3],"#d6604d","#f4a582"))) + 
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(
    #    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Category",
    x = "Exam",
    y = "Proportion of Group A marks"
  ) +
  ggsave("figs/A_breakdown.pdf",width=15,height=10,units="cm",dpi=300)
```









