---
title: "MATH 18/19: Analysis of differences in school exams"
author: "George Kinnear"
date: "28/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(plyr)
library(tidyverse)
library(knitr)
library(BEST) # for Bayesian estimation in place of t-tests
```

## Read data

Read in the `data_math-1819.csv` file which contains the codes assigned to all items, and produce some summaries of this.

```{r read-data, message=FALSE, warning=FALSE, include=FALSE}
source(file = here::here("script_read-data.R"))

School_props = MATH_Group_proportions %>%
  mutate(
    Qual = case_when(
      str_sub(Course,1,7) == "A-Level" ~ "A-Level",
      Course == "SQA AH" ~ "SQA AH",
      Course == "IB HL" ~ "IB HL",
      TRUE ~ "Other"
    )
  ) %>%
  filter(!Qual == "Other") %>%
  ungroup()
```

Sample of the summary data:

```{r summary-papers, echo=FALSE}
School_props %>% 
  head() %>% 
  kable(booktabs = TRUE)
```



# Overall proportions

```{r plot-proportions-abc}
School_props %>%
  group_by(Qual,MATH_Group) %>%
  summarise(
    prop_marks = mean(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    #    Course = paste(Course,"\n(n=",n,")"),
    Qual = (reorder(Qual, gpAmarks, max))
  ) %>%
  ddply(.(Qual), transform, pos = cumsum(prop_marks) - (0.5 * prop_marks)) %>%
  ggplot(aes(x = Qual,
             y = prop_marks,
             fill = fct_rev(MATH_Group))) +
  geom_bar(stat="identity") +
  geom_text(aes(x = Qual, y = pos, label = paste0(sprintf("%.0f",100*prop_marks),"%")),
            size=4) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0,1,0.2)) +
  scale_fill_manual(values = palMATH) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(
    #    title = "Mean proportion of marks in each MATH Group",
    fill = "MATH Group",
    x = "Exam",
    y = "Proportion of marks"
  ) +
  ggsave("figs/ABC_school_means.pdf",width=15,height=10,units="cm",dpi=300)

```


In this version, we sort by the mean % of group A, and show the means with diamonds

```{r plot-props-sorted}
School_props_boxplot_data = School_props %>%
  # reorder the Qual factor by mean % group A
  mutate(
    gpAmarks = if_else(MATH_Group=="A",prop_marks,0),
    Qual = reorder(Qual, gpAmarks, mean)
  )
School_props_boxplot_data %>%
  ggplot(aes(x=1,y=prop_marks, fill=Qual)) +
  geom_boxplot(alpha = 0.3) + 
  stat_summary(fun.y = "mean",
               geom = "point",
               shape = 23,
               size = 3,
               aes(fill = Qual),
               position = position_dodge(0.8)) +
  geom_point(aes(colour=Qual),
             position = position_jitterdodge(),
             size = 1) +
  geom_text(data = School_props_boxplot_data %>%
              group_by(Qual, MATH_Group) %>%
              summarise(prop_marks = mean(prop_marks)),
            aes(label = paste0(100*round(prop_marks,2),"%"),
                y = -0.05),
            position = position_dodge(0.75),
            size = 2.5) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     breaks = seq(0,1,0.1)) +
    scale_fill_manual(values = palMATH) +
    scale_colour_manual(values = palMATH) +
#  scale_fill_brewer(palette = "RdBu", direction = -1) +
#  scale_colour_brewer(palette = "RdBu", direction = -1) +
  theme_light(base_size = 12)+
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    panel.grid.major.x = element_blank(),
    strip.background = element_rect(fill=NA,colour = NA),
    strip.text = element_text(size=10, face="bold", color = "black")
  ) +
  labs(
    #    title = "Proportion of marks in each MATH Group",
    #    subtitle = "Exams are ordered by proportion of Group A marks",
    fill = "Qualification",
    colour = "Qualification",
    x = "MATH Group",
    y = "Proportion of marks"
  ) +
  facet_grid( ~ MATH_Group, switch = "x") +
  ggsave("figs/ABC_school_boxplots.pdf",width=15,height=10,units="cm",dpi=300)
```


# Comparisons

## SQA AH and A-Level

```{r stats-bayes-ah-alevel}
sp = School_props %>%
  filter(MATH_Group == "A") %>%
  select(Qual, prop_marks)

BESTout = BESTmcmc(sp %>% filter(Qual=="SQA AH") %>% select(prop_marks) %>% data.matrix,
                   sp %>% filter(Qual=="A-Level") %>% select(prop_marks) %>% data.matrix)
BESTout
summary(BESTout) %>% kable(booktabs = T)
plot(BESTout)
plot(BESTout, "sd")
plotPostPred(BESTout)
plotAll(BESTout, credMass=0.95, ROPEm=c(-0.1,0.1),
        ROPEeff=c(-0.2,0.2), compValm=0.5)
pairs(BESTout)

```

This shows a mean difference of `r formathdi(summary(BESTout),"muDiff",2)`.

## SQA AH and IB

```{r stats-bayes-ah-ib}
BESToutIB = BESTmcmc(sp %>% filter(Qual=="SQA AH") %>% select(prop_marks) %>% data.matrix,
                   sp %>% filter(Qual=="IB HL") %>% select(prop_marks) %>% data.matrix)
plotAll(BESToutIB, credMass=0.95)
```

This shows a mean difference of `r formathdi(summary(BESToutIB),"muDiff",2)`.

## Overall summary

```{r stats-allquals-means}
School_props %>%
  group_by(Qual, MATH_Group) %>%
  summarise(
    m = mean(prop_marks),
    sd = sd(prop_marks),
    n = n()
  ) %>%
  ungroup() %>%
  arrange(MATH_Group) %>% 
  kable(booktabs = TRUE)
```


## SQA AH and A-Level, Group C

```{r stats-bayes-ah-alevel-gp-c}
school_gpC = School_props %>%
  filter(MATH_Group == "C") %>%
  select(Qual, prop_marks)

BESTout = BESTmcmc(school_gpC %>% filter(Qual=="SQA AH") %>% select(prop_marks) %>% data.matrix,
                   school_gpC %>% filter(Qual=="A-Level") %>% select(prop_marks) %>% data.matrix)
BESTout
summary(BESTout) %>% kable(booktabs = T)
```

This shows a mean difference of `r formathdi(summary(BESTout),"muDiff",2)`.